---
title: "Working with Workflows in PEITHO: Import, Export, and Execution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Workflows in PEITHO: Import, Export, and Execution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to use the PEITHO package to manage, export, import, and execute workflows. The workflow system in PEITHO allows you to define a sequence of steps (such as data fetching, processing, or analysis), save and share workflows as zip files, and re-import or modify them for further use. This guide is aimed at users with basic R knowledge, but no advanced R expertise is required.


# Prerequisites

- Install the PEITHO package from GitHub if not already installed:

  `devtools::install_github("Pandora-IsoMemo/PEITHO")`

- Load the package:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(PEITHO)
```

- For development only, you may use `devtools::load_all()` instead of `library(PEITHO)`.


# Loading the Example Workflow

You can use the example workflow included with PEITHO by passing an empty string as the path:

```{r, echo=TRUE}
my_wf <- new_workflow(workflow_file_paths = workflow_file_paths(path = ""))
```

# Loading a Workflow from a Custom Folder

If you have workflow configuration files (`commands.json`, `inputs.json`, etc.) stored in a local folder, you can load them directly by specifying the folder path:

```{r, eval=FALSE}
my_custom_wf <- new_workflow(
  workflow_file_paths = workflow_file_paths(
    path = file.path("PATH", "TO", "YOUR_CUSTOM_WORKFLOW_FOLDER")
  )
)
```

This approach is useful for:
- Testing custom workflows during development
- Loading workflows from local directories without packaging them as zip files
- Working with multiple workflow variants stored in different folders

# Creating a workflow

A workflow consists of two main configuration files:

## Inputs.json File

The `inputs.json` file defines the input data for your workflow. It's a JSON object with key-value pairs:

```json
{
  "input_name_1": "value or data for first input",
  "input_name_2": "value or data for second input"
}
```

**Example:**
```json
{
  "empire_inputs": "https://en.wikipedia.org/wiki/Roman_Empire\nhttps://en.wikipedia.org/wiki/Achaemenid_Empire",
  "statistics_inputs": "https://en.wikipedia.org/wiki/Coefficient_of_determination"
}
```

## Commands.json File

The `commands.json` file defines the workflow steps. It's a JSON array of command objects, where each object specifies a step in your workflow:

```json
[
  {
    "entry": 1,
    "name": "step_name",
    "label": "Human-readable step label",
    "comments": "Optional comments about this step",
    "command": "function_to_execute",
    "args": "function arguments and references",
    "iteration": "no|yes|auto",
    "samples": 1
  }
]
```

**Field Descriptions:**
- `entry`: Step number (integer)
- `name`: Unique identifier for the step
- `label`: Human-readable description shown in the UI
- `comments`: Optional notes about the step
- `command`: Function/command name to execute
- `args`: Arguments passed to the command. Use `@#*I*#@input_name@#*I*#@` to reference inputs and `@#*L*#@step_name@#*L*#@` to reference outputs from previous steps
- `iteration`: Whether to iterate over results (`no`, `yes`, or `auto`)
- `samples`: Number of samples to process

**Example:**
```json
[
  {
    "entry": 1,
    "name": "Split",
    "label": "Split addresses",
    "comments": "",
    "command": "simple_split",
    "args": "x=@#*I*#@empire_inputs@#*I*#@, split=\\n",
    "iteration": "no",
    "samples": 1
  },
  {
    "entry": 2,
    "name": "get_content",
    "label": "Fetch web content",
    "comments": "",
    "command": "fetch_WebText",
    "args": "url=@#*L*#@Split@#*L*#@",
    "iteration": "auto",
    "samples": 1
  }
]
```


# Exporting a Workflow as a Zip File

You can export the current workflow setup (all relevant files) as a zip archive (here with the `.peitho` extension):

```{r, echo=TRUE}
zipfile_path <- "./examples/my_workflow.peitho"
save_as_zip(my_wf, file = zipfile_path)
```


# Running a Workflow

To execute the workflow, use the `run()` function. You can specify which steps to run (e.g., from step 1 to 5):

```{r, echo=TRUE}
my_run_1 <- run(my_wf, from = 1, to = 5)
```

After running, you can inspect the results:

```{r, echo=TRUE}
length(my_run_1$state$last_result)

PEITHO:::trunc(my_run_1$state$last_result, n_char = 100)
```



# Importing and Modifying a Workflow from a Zip File

To import a workflow from a zip file, first define the directory where the workflow files will be extracted. You may also modify the workflow steps (for example, by editing `commands.json` or `inputs.json` in the extracted folder) before running it.
```{r, echo=TRUE}
zipfile_path <- "./examples/my_workflow.peitho"

my_wf_imported <- PEITHO::import_workflow(
  zipfile = zipfile_path,
  extract_dir = "./examples/peitho_import"
)
```

You can now run the imported (and possibly modified) workflow and inspect the results as before:
```{r, echo=TRUE}
my_run_2 <- run(my_wf_imported, from = 1, to = 4)

length(my_run_2$state$last_result)
PEITHO:::trunc(my_run_2$state$last_result, n_char = 100)
```

# Summary

PEITHO's workflow system allows you to:

- Load workflows from the built-in example or custom folder paths
- Load and execute multi-step workflows from configuration files
- Export workflows as zip files for sharing or backup
- Import and modify workflows from zip files
- Inspect and print results from workflow execution

This makes it easy to share, reuse, and adapt complex analysis pipelines in a reproducible way.
