---
title: "Workflow Quickstart in PEITHO: Load, Run, Import, Export"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow Quickstart in PEITHO: Load, Run, Import, Export}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This quickstart shows the fastest path to run a workflow in PEITHO and share it as a `.peitho` file. You will:

- Load a workflow
- Run selected steps
- Inspect results
- Export and re-import the workflow bundle

For advanced execution topics (resume, robust error handling), see the vignette `peitho_workflow_execution_recovery`. For workflow authoring and editing, see `peitho_workflow_authoring_editing`.


# Prerequisites

- Install the PEITHO package from GitHub if not already installed:

  `devtools::install_github("Pandora-IsoMemo/PEITHO")`

- Load the package:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(PEITHO)
```

- For development only, you may use `devtools::load_all()` instead of `library(PEITHO)`.


# Loading the Example Workflow

You can use the example workflow included with PEITHO by passing an empty string as the path:

```{r, echo=TRUE}
my_wf <- new_workflow(workflow_file_paths = workflow_file_paths(path = ""))
```

You can inspect what was loaded:

```{r, echo=TRUE}
my_wf
```

# Loading a Workflow from a Custom Folder

If you have workflow configuration files (`commands.json`, `inputs.json`, etc.) stored in a local folder, you can load them directly by specifying the folder path:

```{r, eval=FALSE}
# define the workflow
my_custom_wf <- new_workflow(
  workflow_file_paths = workflow_file_paths(
    path = file.path("PATH", "TO", "YOUR_CUSTOM_WORKFLOW_FOLDER")
  )
)

# run the workflow, e.g. steps 2 through 4
my_run <- run(my_custom_wf, from = 2, to = 4)

# check the result of the last step
my_run$state$last_result
```

This approach is useful for:

- Testing custom workflows during development
- Loading workflows from local directories without packaging them as zip files
- Working with multiple workflow variants stored in different folders

# Exporting a Workflow as a Zip File

You can export the current workflow setup (all relevant files) as a zip archive (here with the `.peitho` extension):

```{r, echo=TRUE}
zipfile_path <- "./examples/my_workflow.peitho"
save_as_zip(my_wf, file = zipfile_path)
```


# Running a Workflow

To execute the workflow, use the `run()` function. You can specify which steps to run (e.g., from step 1 to 5):

```{r, echo=TRUE}
my_run_1 <- run(my_wf, from = 1, to = 5)
```

After running, you can inspect the results:

```{r, echo=TRUE}
length(my_run_1$state$last_result)

PEITHO:::trunc(my_run_1$state$last_result, n_char = 100)
```



# Importing and Modifying a Workflow from a Zip File

To import a workflow from a zip file, first define the directory where the workflow files will be extracted. You may also modify the workflow steps (for example, by editing `commands.json` or `inputs.json` in the extracted folder) before running it.
```{r, echo=TRUE}
zipfile_path <- "./examples/my_workflow.peitho"

my_wf_imported <- PEITHO::import_workflow(
  zipfile = zipfile_path,
  extract_dir = "./examples/peitho_import"
)
```

You can now run the imported (and possibly modified) workflow and inspect the results as before:
```{r, echo=TRUE}
my_run_2 <- run(my_wf_imported, from = 1, to = 4)

length(my_run_2$state$last_result)
PEITHO:::trunc(my_run_2$state$last_result, n_char = 100)
```

# Summary

PEITHO's workflow system allows you to:

- Load workflows from the built-in example or custom folder paths
- Load and execute multi-step workflows from configuration files
- Export workflows as zip files for sharing or backup
- Import and modify workflows from zip files
- Inspect and print results from workflow execution

This makes it easy to share, reuse, and adapt complex analysis pipelines in a reproducible way.

# Next Steps

- Advanced execution and recovery (resume runs, error handling): `vignette("peitho_workflow_execution_recovery")`
- Authoring and editing workflows (JSON schema, tags, selectors, update APIs): `vignette("peitho_workflow_authoring_editing")`
