---
title: "Workflow Execution in PEITHO: Errors, Resume, and Recovery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow Execution in PEITHO: Errors, Resume, and Recovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette covers robust workflow execution patterns in PEITHO. It focuses on how to:

- Continue runs after failures
- Control stop-on-error behavior
- Reuse run IDs to resume partially completed runs
- Inspect run metadata for troubleshooting

If you are new to PEITHO workflows, start with `vignette("peitho_workflow_basics")`.

# When to Use This Vignette

Use this vignette when your workflow runs are long-running, error-prone, or need restart/resume behavior.

# Prerequisites

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(PEITHO)
```

# Load a Workflow

```{r, echo=TRUE}
wf <- new_workflow(workflow_file_paths = workflow_file_paths(path = ""))
```

# Basic Run and Run Metadata

```{r, echo=TRUE}
run_1 <- run(wf, from = 1, to = 3)
run_1$run_id
```

The `run_id` uniquely identifies a workflow execution and is used for resume operations.

# Error Handling Strategy

By default, `run()` uses `stop_on_error = TRUE`. You can set `stop_on_error = FALSE` to continue execution and inspect errors afterward.

```{r, eval=FALSE}
run_nonstop <- run(
  wf,
  from = 1,
  to = 5,
  stop_on_error = FALSE
)

run_nonstop$errors
```

Use this mode when you want a full pass and post-run diagnosis, rather than fail-fast behavior.

# Resume a Run

If a run stops midway, resume using the same `run_id`.

```{r, eval=FALSE}
# First run (example may stop due to an error):
first_try <- run(wf, from = 1, to = 5)

# Resume from saved run ID:
resumed <- run(
  wf,
  from = 1,
  to = 5,
  resume_run_id = first_try$run_id
)
```

`resume_run_id` tells PEITHO to continue from the first unfinished step/iteration recorded for that run.

# Practical Recovery Workflow

A common pattern for robust execution:

1. Run with `stop_on_error = TRUE` during normal usage.
2. Capture `run_id` when an interruption/error occurs.
3. Fix the issue (input values, custom function, environment, etc.).
4. Re-run with `resume_run_id` to avoid repeating finished work.

```{r, eval=FALSE}
safe_run <- run(wf, from = 1, to = 5, stop_on_error = TRUE)

# ... investigate/fix root cause ...

safe_run_resumed <- run(
  wf,
  from = 1,
  to = 5,
  resume_run_id = safe_run$run_id,
  stop_on_error = TRUE
)
```

# Troubleshooting Checklist

- Confirm workflow files exist and are readable.
- Ensure command functions are available in the expected environment.
- Verify inputs referenced in command args actually exist in `input_list`.
- Check the run object (`$errors`, `$state`, `$run_id`) before retrying.

# Summary

You now have a robust execution workflow:

- Run with explicit error strategy
- Capture and reuse `run_id`
- Resume incomplete runs without reprocessing completed steps

For schema details and editing workflows, continue with `vignette("peitho_workflow_authoring_editing")`.

# Related Vignettes

- Quickstart (load, run, import/export): `vignette("peitho_workflow_basics")`
- Authoring and editing (files, tags, selectors, update APIs): `vignette("peitho_workflow_authoring_editing")`
